---
title: "How to Track one's life."
description: |
  Creating a workflow for seamless, effortless collection and visualization of daily activities.
  Timetracking part 3 of 5
author: 
 - name: Vadym Yudenko
   affiliation: American University Kyiv
   affiliation-url: https://auk.edu.ua/en/
date: 2023-07-14
draft: true 
categories:
  - timetracking
  - self-surveillance
---

::: {.callout-tip title="Part 1. Every living hour" collapse="true" icon="false" style="text-align: center;"}
[Part 1](https://vyudenko.blog/posts/time-tracking-1) of 5 of time tracking series. It serves as an introduction to the project and it's rationale. Other parts will go in depth and answer specific questions.
:::

::: {.callout-tip title="Part 2. Why track your daily life?" collapse="true" icon="false" style="text-align: center;"}
[Part 2](https://vyudenko.blog/posts/time-tracking-2) explains timetracking and goes over some of the reasons why you should or shouldn't track your daily activities. Who will benefit from using it and who won't.
:::

::: {.callout-tip title="Part 3. How to track your daily life?" collapse="false" icon="false" style="text-align: center;"}
In this section you'll read about different approaches to tracking one's time based on the goals you have in mind. Not everyone should do it the same way!
:::

::: {.callout-tip title="Part 4. Answering questions with data" collapse="true" icon="false" style="text-align: center;"}
[Part 4](https://vyudenko.blog/posts/time-tracking-4) provides examples of what kind of data-driven answers you can get by tracking your time.
:::

::: {.callout-tip title="Part 5. Other" collapse="true" icon="false" style="text-align: center;"}
[Part 5](https://vyudenko.blog/posts/time-tracking-5) looks at other projects....
:::


## Methodology (how have I approached this)

::: {.callout-tip title="Setup" collapse="false" icon="false" style="text-align: center;"}
1.  Come up with categories
2.  Create a spreadsheet
3.  Pre-fill routines
4.  Track
:::

This goes along with my thinking about my life as multiple spheres. In order to lead a fulfilling, meaningful and balanced life I have to be successful in multiple of these spheres.

[Read about it in my article](https://vyudenko.blog/posts/time-tracking/)

## Categories

Before you jump into tracking, think carefully what it is you're tracking. What do you want to do with this data? It is a long-term commitment and you don't want to be stuck with a system that doesn't work for you. You can always change it, but it is better to think about it beforehand.

Myself, a student at the time, had not much going in my life, except for studying in university, learning Japanese, reading books and doing sports. Added these to the basic categories of Sleep, Food and Transport. Last was the inclusion of Waste of Time and Miscellaneous.

Main categories:

### Study/Work

This category includes everything that is related to my studies and work. This includes classes, homework, studying, research, writing, reading, etc. Later on I started counting my non-university studies as study as well. So one can think of it as 'brain activity'.

### Japanese

I have started to learn Japanese in early 2019 as a hobby. It took time so I decided to track it separately from Study. Later on I became more serious about it and in order to prepare for the language proficiency test I calculated the amount of time I need to spend on it by December 2020 to pass.

### Reading

I used to read quite a lot before uni and before starting to learn Japanese. This category was meant for all the non-uni, non-Japanese reading - reading for pleasure.

### Planning

This category is for journaling, planning, thinking about my life, etc. It is a very important category for me, as it helps me to stay on track and not get lost in the daily routine. I have intended in as a writing practice - writing in a diary, but I ended up mostly thinking to myself.

### Friends and Family

I have been spending less and less time with my family, relying on phone calls and messages. So, this category was included in part to track how much time I spend with my family and friends and in part to motivate me to spend more time with them. I created a goal of spending at least 6 hours a week with them.

### Sports

Healthy body, healthy mind. I believe in this, so tracked my sports activities. I have included all the sports activities, including walking, running and playing frisbee. This category, like Japanese had a goal - to spend on average 1 hour a day on sports.

### Sleep

Sleep includes night sleep, occasional naps and rests. In other words, the time for resting and recharging.

### Food

This category wasn't meant to track the length of my meals, which are quite short, but rather capture eating patterns. How many meals a day? What time?

### Transport

This category is meant to capture the time I spend on going somewhere. Mostly getting around the city and travelling. It can be thought of as one of Miscellaneous category, which I explain in the end.

Aster Hu's says about this category....

### Miscellaneous

The most mysterious category. Everything that doesn't fit into the above categories, goes here. It is not, unproductive time, per se, but it isn't something I should minimize. This category includes household chores, doctor visits, haircuts, shopping, some other 'I have to do this things'.

A lot of the times this category overlaps heavily with other categories. For example, I can be listening to a podcast while doing chores, or I can be reading a book while waiting for an appointment.

### Waste of Time

If you find yourself thinking: 'I wish I hadn't spend this time this way...', then this is a candidate for this category. It is a category for unproductive time, especially if you think you could've spent it better. It is easier to understand with an example. You came home from work, you're tired, you want to relax, but you end up watching YouTube for 3 hours straight. You feel bad about it, since you also had to walk your dog, do the laundry, or you planned to read that book this evening. The feeling of regret goes hand in hand with this category for me.

### Quiz

These categories are not set in stone, but rather a guide for you to easily categorize your activities. Consider the following situations you will encounter as you track your time.

You are on a bus on your way to work. You are reading a book. How do you categorize this time?

```{r, results='asis'}
#| warning: false
#| echo: false
#| message: false
require("checkdown")
check_question(c(""), options = c("Reading", "Transport", "Study", "Miscellaneous (Other)", "Waste of time"), type = "radio", wrong = "Sure!", random_answer_order = TRUE)
```

<hr>

You woke up at 3am and can't fall asleep, you just lie there turning from side to side and it goes on for an hour. Until you decide to get up and do something else, like watching news. How do you categorize this time?

```{r, results='asis'}
#| warning: false
#| echo: false
#| message: false
require("checkdown")
check_question(c(""), options = c("Sleeping", "Miscellaneous (Other)", "Waste of time"), type = "radio", right = "Of course!", wrong = "Sure!", random_answer_order = TRUE)
```

<hr>

You are cooking together with your partner or a friend. You are talking about your day, about your plans, about your life. How do you categorize this time?

```{r, results='asis'}
#| warning: false
#| echo: false
#| message: false
require("checkdown")
check_question(c(""), options = c("Food", "Family and friends (Communication)", "Planning", "Miscellaneous (Other)"), type = "radio", wrong = "Sure!", random_answer_order = TRUE)
```

## Spreadsheet

Now that we have categories and hopefully things we want to track in hours. We can proceed to setting up tracking. The idea of doing it in Excel comes from [Coffee Break's video](https://www.youtube.com/watch?v=Owgxw8d4L84) I watched a while ago.

We are using the following setup: Rows - Time; Columns - Days; Cells - Categories.

Time intervals can be a bit tricky, since the more precise you want to be, the more time it takes to track. The opposite is also true - the larger the intervals, the less accurate the data. The choice of time intervals of 20 mins was because at the time it seemed to strike a balance between precision and ease of tracking. Using 10 or 15 mins intervals seemed like an overkill[^1], 30 minutes would be ideal, as it strikes the balance between the effort and precision. I've ultimately settled on 20 minutes.

[^1]: I do realize that this whole setup can be viewed as an overkill by some, but this seemed okay at the time. I got used to it, and now it takes minimum effort.

Somewhere in your file create an array of *categories*. Then create a dropdown list for each cell. Lastly create a conditional formatting rule, which will color the cell depending on the category. After you are done, your spreadsheet should look something like this: <br><br>

::: column-page-left
<iframe src="https://docs.google.com/spreadsheets/d/1HYKMcbxnRHw1CLmeIymic0JXURoay6vB/edit?usp=sharing&amp;ouid=106672896824390417572&amp;rtpof=true&amp;sd=true" width="1200" height="800">

</iframe>
:::

## Pre-fill routines

One thing this workflow should achieve is be easy to use and take as little time as possible. We have already created a dropdown list for each cell, settled on the time intervals, now we can pre-fill the routines. For example, I have a morning routine, which I do every day. I can pre-fill it for the whole week. Or I can pre-fill my university schedule for the whole semester. Ctrl+C and Ctrl+V are your friends here.

The only thing you have to do is to fill the gaps and change pre-filled routines if needed.

## Track

Now that you have your spreadsheet set up, you can start tracking! I suggest you install the Google Sheets app on your phone, so that you can track on the go. And make sure to enable Sync or manually save a copy of the files not to lose your data!

When tracking remember your final goal and take it easy. This shouldn't be a burden, it, ideally, should become a part of your routine. There is no need to log the data every hour, you can do it every 3-4 hours, when you have a break, on your way home or before going to bed.

Make sure you always keep the cognitive load low. If you often have to think about what category to choose, then you have too many categories. Remember that it is a long-term project, better develop a habit and then it will become easier to track. But keep cognitive load low!

## Conclusion

There are many ways to track your time, which are all good in their own way. I have tried to explain my way of tracking time, which I have been using for the past 2 years. I hope you found it useful and will try it out. 

> Like a Calendar app for masochists.

Read next:

![](https://vyudenko.blog/posts/time-tracking-4
