---
title: "Other resource"
description: |
 .....
  Timetracking part 5 of 5
author: 
 - name: Vadym Yudenko
   affiliation: American University Kyiv
   affiliation-url: https://auk.edu.ua/en/
date: 2023-07-14
draft: true 
categories:
  - timetracking
  - self-surveillance
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load libraries}
# devtools::install_github("liamgilbey/ggwaffle")
# library(ggwaffle)
library(stargazer)
library(tidyverse)
library(ggExtra)
library(ggtext)
library(ggthemes)
library(gganimate)
library(ggnewscale)
library(showtext)
library(grid)
library(transformr)
library(corrr)
library(corrplot)
library(lubridate)
library(reshape2)
library(plotly)
library(scales)
library(ggiraph)
library(glue)
library(knitr)
library(fontawesome)
library(RColorBrewer)
library(DT)
library(patchwork)
```

```{r set theme}
font_add_google("Fira Sans", "firasans")
font_add_google("IBM Plex Mono", "ibmplex")
# showtext::showtext_auto()

blog_theme <-
   theme_minimal(18, base_family = "IBM Plex Mono") +
   theme(
     plot.background = element_rect(fill = "white", color = NA),
     plot.title.position = "plot",
     plot.title = element_text(size = 24, margin = margin(b = 1, unit = "line")),
     legend.position = c(0, 1),
     legend.direction = "horizontal",
     legend.justification = c(0, 1),
     legend.title.align = 1,
     axis.title.y = element_text(hjust = 0, size = 10),
     axis.title.x = element_text(hjust = 0, size = 10),
     panel.grid.major = element_line(color = "#d3d9db"),
     panel.grid.minor = element_blank()
   )
 
theme_set(blog_theme)
```

```{r load data}
data_long <- read.csv("data/data.csv") %>%
  mutate(
    datetime = as.POSIXct(datetime, format="%m/%d/%Y %H:%M", tz="GMT+3"),
    day_of_week = weekdays(datetime),
    time_of_day = case_when(
      hour(datetime) < 12 ~ "Morning",
      hour(datetime) < 18 ~ "Day",
      TRUE ~ "Night"
    ),
    time_column = if_else(substr(as.character(datetime), 12, 16) == "", "00:00", substr(as.character(datetime), 12, 16)),
    category5 = case_when(
      activity %in% c("Sleep", "Sport", "Communication", "Japanese") ~ activity,
      activity %in% c("Study", "Work", "TM", "Planning", "Reading") ~ "Study/Work",
      activity %in% c("Transportation", "Misc", "Waste", "Food") ~ "Rest"
    ),
    hour = hour(datetime),
    day = wday(datetime, label = TRUE, abbr = TRUE, week_start = 1),
    month = month(datetime, label = TRUE, abbr = FALSE),
    date_char = as.character(datetime)
  )

activity_colors <- c(
  "Sleep" = "#1E90FF",  # DodgerBlue
  "Sport" = "#32CD32",  # LimeGreen
  "Study" = "#DAA520",  # Goldenrod
  "Communication" = "#87CEEB",  # SkyBlue
  "Food" = "#FFD700",  # Gold
  "Waste" = "black",
  "Reading" = "#FFA07A",  # LightSalmon
  "Japanese" = "#FF69B4",  # HotPink
  "Misc" = "#8FBC8F",  # DarkSeaGreen
  "TM" = "#9370DB",  # MediumPurple
  "Transportation" = "#A9A9A9"  # DarkGray
)

time_ticks06 <- c(
    "06:00", "06:20", "06:40", "07:00", "07:20", "07:40", "08:00", "08:20", "08:40", "09:00",
    "09:20", "09:40", "10:00", "10:20", "10:40", "11:00", "11:20", "11:40", "12:00", "12:20",
    "12:40", "13:00", "13:20", "13:40", "14:00", "14:20", "14:40", "15:00", "15:20", "15:40",
    "16:00", "16:20", "16:40", "17:00", "17:20", "17:40", "18:00", "18:20", "18:40", "19:00",
    "19:20", "19:40", "20:00", "20:20", "20:40", "21:00", "21:20", "21:40", "22:00", "22:20",
    "22:40", "23:00", "23:20", "23:40", "00:00", "00:20", "00:40", "01:00", "01:20", "01:40",
    "02:00", "02:20", "02:40", "03:00", "03:20", "03:40", "04:00", "04:20", "04:40", "05:00",
    "05:20", "05:40"
)


time_ticks15 <- c("15:00", "15:20", "15:40", "16:00", "16:20", "16:40", "17:00",
                "17:20", "17:40", "18:00", "18:20", "18:40", "19:00", "19:20",
                "19:40", "20:00", "20:20", "20:40", "21:00", "21:20", "21:40",
                "22:00", "22:20", "22:40", "23:00", "23:20", "23:40",
                "00:00", "00:20", "00:40", "01:00", "01:20", "01:40",
                "02:00", "02:20", "02:40", "03:00", "03:20", "03:40",
                "04:00", "04:20", "04:40", "05:00", "05:20", "05:40",
                "06:00", "06:20", "06:40", "07:00", "07:20", "07:40",
                "08:00", "08:20", "08:40", "09:00", "09:20", "09:40",
                "10:00", "10:20", "10:40", "11:00", "11:20", "11:40",
                "12:00", "12:20", "12:40", "13:00", "13:20", "13:40", "14:00", "14:20", "14:40")
```

```{r atus}
#sources: https://rpubs.com/Ryder/555470
# atus.sum <- read.delim('data/atussum_0322.dat', sep=",")
# atus.who <- read.delim('data/atuswho_0322.dat', sep=",")
# atus.act <- read.delim('data/atusact_0322.dat', sep=",")
# atus.rostec <- read.delim('data/atusrostec_1122.dat', sep=",")
# atus.rost <- read.delim('data/atusrost_0322.dat', sep=",")
# atus.case <- read.delim('data/atuscase_0522.dat', sep=",")
# # Joining these by TUCASEID
# atus <- atus.sum %>%
#   left_join(atus.who, by = "TUCASEID") %>%
#   left_join(atus.act, by = "TUCASEID") %>%
#   left_join(atus.rostec, by = "TUCASEID") %>%
#   left_join(atus.rost, by = "TUCASEID") %>%
#   left_join(atus.case, by = "TUCASEID")
```

```{r}

```


::: {.callout-tip title="Part 1. Every living hour" collapse="true" icon="false" style="text-align: center;"}
[Part 1](https://vyudenko.blog/posts/time-tracking-1) of 5 of time tracking series. It serves as an introduction to the project and it's rationale. Other parts will go in depth and answer specific questions.
:::

::: {.callout-tip title="Part 2. Why track your daily life?" collapse="true" icon="false" style="text-align: center;"}
[Part 2](https://vyudenko.blog/posts/time-tracking-2) explains timetracking and goes over some of the reasons why you should or shouldn't track your daily activities. Who will benefit from using it and who won't.
:::

::: {.callout-tip title="Part 3. How to track your daily life?" collapse="true" icon="false" style="text-align: center;"}
[Part 3](https://vyudenko.blog/posts/time-tracking-3) lays out the approach and its set up. From choosing the categories to the actual tracking.
:::

::: {.callout-tip title="Part 4. Answering questions with data" collapse="true" icon="false" style="text-align: center;"}
[Part 4](https://vyudenko.blog/posts/time-tracking-4) provides examples of what kind of data-driven answers you can get by tracking your time.
:::

::: {.callout-tip title="Part 5. Other" collapse="false" icon="false" style="text-align: center;"}
In this part we look at other projects....
:::

Sources: https://flowingdata.com/2022/09/01/looking-for-meaning-in-the-everyday/ https://flowingdata.com/2021/11/23/most-common-daily-schedules-for-different-groups/ 
https://flowingdata.com/2017/10/19/american-daily-routine/ 
https://flowingdata.com/2023/07/25/when-is-dinner-by-state/ 
https://flowingdata.com/2011/03/20/daily-schedule-of-benjamin-franklin/ https://flowingdata.com/2019/07/16/how-the-american-work-day-changed-in-15-years/ https://lifehacker.com/how-to-track-everything-in-your-life-without-going-craz-1466537828 
https://en.wikipedia.org/wiki/Lifelog#cite_note-1 
https://www.nytimes.com/2007/10/29/us/29shields.html 
https://flowingdata.com/2014/09/04/when-people-work-by-job-category/ 
https://flowingdata.com/2019/04/02/data-comic-shows-an-average-american-day/ 
https://flowingdata.com/2021/08/19/daily-routine-2020/ 
https://flowingdata.com/2015/11/30/most-common-use-of-time-by-age-and-sex/
