---
title: Counting Jellybeans
description: |
  Visualizing one's life.
author: 
 - name: Vadym Yudenko
   affiliation: American University Kyiv
   affiliatiom-url: https://auk.edu.ua/en/
citation: true
date: 2023-07-14
draft: true 
categories:
  - timetracking
  - data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
# devtools::install_github("liamgilbey/ggwaffle")
library(ggwaffle)
library(stargazer)
library(tidyverse)
library(ggExtra)
library(ggthemes)
library(corrr)
library(corrplot)
library(lubridate)
library(reshape2)
library(plotly)
library(scales)
```

# Introduction

One summer day in 2019 unburdened by university studies and need for work, I spend days doing a whole lot of nothing.

After getting up closer to 12 am for a seventh day in a row - and feeling so miserable and tired. I thought that I am wasting too much time in the morning and am more productive in the evening. But is it so? How much hours exactly am I wasting in the morning?

https://youtu.be/BOksW_NabEk

Counting jellybeans

{{< video https://youtu.be/iJNs9nxoQaw >}}

```{r}
data <- read.csv("KSE_life.csv")

data_long <- gather(data, day, activity, -X)

data_long$day <- sub("^X", "", data_long$day)

data_long$datetime <- as.POSIXct(paste(data_long$day, data_long$X), format="%m.%d.%Y %H:%M")

data_long <- data_long[, !(names(data_long) %in% c("day", "X"))]

data_long$day_of_week <- weekdays(data_long$datetime)
data_long$time_of_day <- ifelse(as.numeric(format(data_long$datetime, "%H")) < 12, "Morning",
                         ifelse(as.numeric(format(data_long$datetime, "%H")) < 18, "Day", "Night"))

data_long$time_column <- substr(as.character(data_long$datetime), 12, 16)
data_long$time_column <- ifelse(data_long$time_column == "", "00:00", data_long$time_column)

data_long <- data_long %>% 
  mutate(category5 = case_when(
    activity %in% c("Sleep", "Sport", "Communication") ~ activity,
    activity %in% c("Study", "Work", "TM", "Planning", "Reading") ~ "Study/Work",
    activity %in% c("Transport", "Misc", "Waste", "Food") ~ "Rest"
  ))
```

```{r}
p <- data_long %>% 
  mutate(date = as.Date(datetime)) %>% 
  ggplot(aes(date, time_column, fill = category5)) +
  geom_tile(
    aes(
      text = str_replace_all(paste0(
        "<b>", time_column, "</b><br>",
        "Activity: ", category5, "<br>"), "NA", "missing")
      )
    ) +
  labs(x = NULL, y = NULL, title = "Daily Behavioral Patterns") +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 14)
    )

ggplotly(p, tooltip = "text") %>%
  layout(
    xaxis = list(
      type = "date",
      tickformat = "%b %-d"
    ),
    yaxis = list(fixedrange = TRUE),
    font = list(family = "IBM Plex Sans Condensed"),
    hoverlabel = list(font = list(family = "IBM Plex Sans Condensed"), align = "left"),
    # Add date range slider
    xaxis2 = list(
      type = "date",
      tickformat = "%b %-d",
      overlaying = "x",
      side = "bottom"
    ),
    margin = list(l = 50, r = 50, t = 20, b = 50),
    updatemenus = list(
      list(
        type = "buttons",
        showactive = FALSE,
        buttons = list(
          list(
            step = "all",
            label = "Reset",
            method = "relayout",
            args = list("xaxis.range", list(NULL, NULL))
          )
        )
      )
    )
  ) %>%
  #rangeslider(start = data_long$datetime[287], end = data_long$datetime[1398]) %>%
    plotly::config(
      displayModeBar = TRUE,
      displaylogo = FALSE,
      modeBarButtonsToRemove = c("zoom2d", "pan2d", "zoomIn2d", "zoomOut2d",
                                 "autoScale2d", "toggleSpikelines", "hoverClosestCartesian",
                                 "hoverCompareCartesian", "toImage")
      )

```



# Activities

```{r}
df_activity_hours <- data_long %>%
  mutate(date = as.Date(datetime)) %>%
  group_by(date, activity) %>%
  summarise(hours = n() * (20/60)) %>%
  ungroup()

# Plot the data
p <- ggplot(df_activity_hours[287:1398,], aes(x = date, y = hours, color = activity)) +
  geom_line(size = 0.75) +  # Customize the line thickness here
  labs(x = "Date", y = "Hours", color = "Activity") +
  theme_minimal()

# Convert it to a plotly object
p <- ggplotly(p, tooltip = c("x", "y", "color"))

p
```



## Sleeping 

```{r}

```



# Recomendations





::: {.scatter-wide .l-page}
```{r, layout="l-page", out.height="400px", eval=T}
subplot(scatter, titleX = TRUE, shareY = TRUE) %>%
  highlight(
    on = "plotly_hover",
    off = "plotly_doubleclick",
    opacityDim = 0.25,
    selectize = TRUE
    )
```
:::

```{r}
p <- mun_acs_calc %>% 
  ggplot(aes(poverty_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Families below 200% poverty line: ", percent(poverty_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.05) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.5, by = 0.05)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Families below 200% of poverty line",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p_plotly <- ggplotly_config(p)

p2 <- mun_acs_calc %>%
  ggplot(aes(crowded_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Overcrowded households: ", percent(crowded_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.05) + 
  scale_x_continuous(label = label_percent(1), breaks = c(0, 0.05, 0.1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.5, by = 0.05)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Overcrowded households",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p2_plotly <- ggplotly_config(p2)

p3 <- mun_acs_calc %>%
  ggplot(aes(black_lat_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Black or Latino population: ", percent(black_lat_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.05) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.5, by = 0.05)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Black or Latino population",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p3_plotly <- ggplotly_config(p3)
```

::: scatter-long
```{r eval=T}
p_plotly
```

```{r eval=T}
p2_plotly
```

```{r eval=T}
p3_plotly
```
:::
