---
title: "What can one learn from analysing your daily life"
description: |
  Applying statistics tools to study daily behavioral patterns.
  Timetracking part 4 of 5
author: 
 - name: Vadym Yudenko
   affiliation: American University Kyiv
   affiliation-url: https://auk.edu.ua/en/
date: 2023-07-14
draft: true 
categories:
  - timetracking
  - self-surveillance
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load libraries}
# devtools::install_github("liamgilbey/ggwaffle")
# library(ggwaffle)
library(stargazer)
library(tidyverse)
library(ggExtra)
library(ggtext)
library(ggthemes)
library(gganimate)
library(ggnewscale)
library(showtext)
library(grid)
library(transformr)
library(corrr)
library(corrplot)
library(lubridate)
library(reshape2)
library(plotly)
library(scales)
library(ggiraph)
library(glue)
library(knitr)
library(fontawesome)
library(RColorBrewer)
library(DT)
library(patchwork)
```

```{r set theme}
font_add_google("Fira Sans", "firasans")
font_add_google("IBM Plex Mono", "ibmplex")
# showtext::showtext_auto()

blog_theme <-
    theme_minimal(18, base_family = "IBM Plex Mono") +
    theme(
      plot.background = element_rect(fill = "white", color = NA),
      plot.title.position = "plot",
      plot.title = element_text(size = 24, margin = margin(b = 1, unit = "line")),
      legend.position = c(0, 1),
      legend.direction = "horizontal",
      legend.justification = c(0, 1),
      legend.title.align = 1,
      axis.title.y = element_text(hjust = 0, size = 10),
      axis.title.x = element_text(hjust = 0, size = 10),
      panel.grid.major = element_line(color = "#d3d9db"),
      panel.grid.minor = element_blank()
    )
  
theme_set(blog_theme)
```

```{r load data}
data_long <- read.csv("data.csv") %>%
  mutate(
    datetime = as.POSIXct(datetime, format="%m/%d/%Y %H:%M", tz="GMT+3"),
    day_of_week = weekdays(datetime),
    time_of_day = case_when(
      hour(datetime) < 12 ~ "Morning",
      hour(datetime) < 18 ~ "Day",
      TRUE ~ "Night"
    ),
    time_column = if_else(substr(as.character(datetime), 12, 16) == "", "00:00", substr(as.character(datetime), 12, 16)),
    category5 = case_when(
      activity %in% c("Sleep", "Sport", "Communication", "Japanese") ~ activity,
      activity %in% c("Study", "Work", "TM", "Planning", "Reading") ~ "Study/Work",
      activity %in% c("Transportation", "Misc", "Waste", "Food") ~ "Rest"
    ),
    hour = hour(datetime),
    day = wday(datetime, label = TRUE, abbr = TRUE, week_start = 1),
    month = month(datetime, label = TRUE, abbr = FALSE),
    date_char = as.character(datetime),
    
    sleep = ifelse(activity == "Sleep", 1, 0),
    sport = ifelse(activity == "Sport", 1, 0),
    study = ifelse(activity == "Study", 1, 0),
    food = ifelse(activity == "Food", 1, 0),
    comm = ifelse(activity == "Communication", 1, 0),
    misc = ifelse(activity == "Misc", 1, 0),
    waste = ifelse(activity == "Waste", 1, 0),
    transport = ifelse(activity == "Transportation", 1, 0),
    jap = ifelse(activity == "Japanese", 1, 0),
    plan = ifelse(activity == "Planning", 1, 0)
  )

activity_colors <- c(
  "Sleep" = "#1E90FF",  # DodgerBlue
  "Sport" = "#32CD32",  # LimeGreen
  "Study" = "#DAA520",  # Goldenrod
  "Communication" = "#87CEEB",  # SkyBlue
  "Food" = "#FFD700",  # Gold
  "Waste" = "black",
  "Reading" = "#FFA07A",  # LightSalmon
  "Japanese" = "#FF69B4",  # HotPink
  "Misc" = "#8FBC8F",  # DarkSeaGreen
  "TM" = "#9370DB",  # MediumPurple
  "Transportation" = "#A9A9A9"  # DarkGray
)

time_ticks06 <- c(
    "06:00", "06:20", "06:40", "07:00", "07:20", "07:40", "08:00", "08:20", "08:40", "09:00",
    "09:20", "09:40", "10:00", "10:20", "10:40", "11:00", "11:20", "11:40", "12:00", "12:20",
    "12:40", "13:00", "13:20", "13:40", "14:00", "14:20", "14:40", "15:00", "15:20", "15:40",
    "16:00", "16:20", "16:40", "17:00", "17:20", "17:40", "18:00", "18:20", "18:40", "19:00",
    "19:20", "19:40", "20:00", "20:20", "20:40", "21:00", "21:20", "21:40", "22:00", "22:20",
    "22:40", "23:00", "23:20", "23:40", "00:00", "00:20", "00:40", "01:00", "01:20", "01:40",
    "02:00", "02:20", "02:40", "03:00", "03:20", "03:40", "04:00", "04:20", "04:40", "05:00",
    "05:20", "05:40"
)


time_ticks15 <- c("15:00", "15:20", "15:40", "16:00", "16:20", "16:40", "17:00",
                "17:20", "17:40", "18:00", "18:20", "18:40", "19:00", "19:20",
                "19:40", "20:00", "20:20", "20:40", "21:00", "21:20", "21:40",
                "22:00", "22:20", "22:40", "23:00", "23:20", "23:40",
                "00:00", "00:20", "00:40", "01:00", "01:20", "01:40",
                "02:00", "02:20", "02:40", "03:00", "03:20", "03:40",
                "04:00", "04:20", "04:40", "05:00", "05:20", "05:40",
                "06:00", "06:20", "06:40", "07:00", "07:20", "07:40",
                "08:00", "08:20", "08:40", "09:00", "09:20", "09:40",
                "10:00", "10:20", "10:40", "11:00", "11:20", "11:40",
                "12:00", "12:20", "12:40", "13:00", "13:20", "13:40", "14:00", "14:20", "14:40")
```

::: {.callout-tip title="Part 1. Every living hour" collapse="true" icon="false" style="text-align: center;"}
[Part 1](https://vyudenko.blog/posts/time-tracking-1) of 5 of time tracking series. It serves as an introduction to the project and it's rationale. Other parts will go in depth and answer specific questions.
:::

::: {.callout-tip title="Part 2. Why track your daily life?" collapse="true" icon="false" style="text-align: center;"}
[Part 2](https://vyudenko.blog/posts/time-tracking-2) explains timetracking and goes over some of the reasons why you should or shouldn't track your daily activities. Who will benefit from using it and who won't.
:::

::: {.callout-tip title="Part 3. How to track your daily life?" collapse="true" icon="false" style="text-align: center;"}
[Part 3](https://vyudenko.blog/posts/time-tracking-3) lays out the approach and its set up. From choosing the categories to the actual tracking.
:::

::: {.callout-tip title="Part 4. Answering questions with data" collapse="true" icon="false" style="text-align: center;"}
In this part we provide examples of what kind of data-driven answers you can get by tracking your time.
:::

::: {.callout-tip title="Part 5. Other" collapse="true" icon="false" style="text-align: center;"}
[Part 5](https://vyudenko.blog/posts/time-tracking-5) looks at other projects....
:::

Non-linear regressions

categorical-categorical \[check CHATGDP\]

```{r}
library(networkD3)
# Load energy projection data
URL <- paste0('https://cdn.rawgit.com/christophergandrud/networkD3/',
              'master/JSONdata/energy.json')
energy <- jsonlite::fromJSON(URL)



class(energy)


nodes <- data.frame(name = c("Sleep", "Sport", "Study", "Communication", "Food", "Waste", "Reading", "Japanese", "Misc", "TM", "Transportation"))

links <- data.frame(source = c(1, 0, 1, 1, 0, 1, 1, 1, 1, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                    target = c(0, 3, 4, 5, 6, 7, 8, 9, 10, 11, 3, 4, 5, 6, 7, 8, 9, 10, 11, 0),
                    value = c(8, 2, 2, 2, 1, 1, 1, 1, 1, 1, 2, 2, 1, 1, 1, 1, 1, 1, 1, 8))

dt <- list(nodes = nodes, links = links)

# Colour Nodes
sankeyNetwork(Links = dt$links, Nodes = dt$nodes, Source = 'source',
              Target = 'target', NodeID = 'name', fontSize = 12, nodeWidth = 30)
 
```

```{r}
library(ggalluvial)

# Step 1: Create a lagged dataset to find subsequent activities
data_long_lagged <- data_long %>%
  mutate(activity_lag1 = lead(activity, 1),
         activity_lag2 = lead(activity, 2),
         activity_lag3 = lead(activity, 3)) %>% 
    filter(activity != activity_lag1 & activity != activity_lag2) # Filter to morning activities

# Step 2: Filter to rows where the activity is "Sleep"
transitions <- data_long_lagged %>%
  filter(activity == "Sleep") %>%
  select(activity, activity_lag1, activity_lag2, activity_lag3)
# Step 3: Create an alluvial plot
ggplot(transitions,
       aes(axis1 = activity, axis2 = activity_lag1, axis3 = activity_lag2, axis4 = activity_lag3)) +
  geom_alluvium(aes(fill = activity), width = 0, decreasing = FALSE,
                   curve_type = "sigmoid") +
  geom_stratum(aes(color = activity)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Wake up", "20 min", "40 min", "60 min"),
                   expand = c(0, 0.05, 0.1, 0.15)) +
  labs(title = "Activity Transitions after Sleep", x = "", y = "Frequency") +
  theme(legend.position = "none")

```




```{r}
df_activity_hours <- data_long %>%
   mutate(date = as.Date(datetime)) %>%
   group_by(date, activity) %>%
   summarise(hours = n() * (20/60)) %>%
   ungroup() 

p <- ggplot(df_activity_hours, aes(x = date)) +
  geom_line(aes(y = hours, color = activity), size = 0.5) +  # Original data
  scale_x_date(limits = c(min(df_activity_hours$date), max(df_activity_hours$date)), 
                   expand = c(0, 0),
                   date_breaks = "1 month",  # Display every month on the x-axis
                   date_labels = "%B") +  # Format the date labels as "Month Year"
  labs(title = "Time spent on activities",
       x = NULL,
       y = "",
       color = "Activity") +  # Labels
  theme(legend.position = "none",
        plot.title = element_text(family = "IBM Plex Mono", hjust = 0.5, size = 18),
        axis.text = element_text(family = "IBM Plex Mono"),
        axis.title = element_text(family = "IBM Plex Mono"),
        axis.text.x = element_text(size = 7))

ggplotly(p)
```

```{r graveyard, echo=FALSE}
# Balanced life
# 
# On the surface it might look like a a purely time tracking task that is no use in planning.
# 
# Yet, after some time one can do the following….
# 
# Think how much time do you have to spend communicating to feel fulfilled, how much time do you wish to spend doing sports? After all both sports and communications are important for your well being. Now, seeing how busy you are, how much time on average do you think you can feasibly speand on these activities.
# 
# (my thinking behind importance of having a balanced life) - you can't just invest all the time into work or family. - there has to be some sort of balance, which can be quantified with some precision
# 
# Reflection
# 
# (thinking behind importance of looking back) - to think through how have you spent one's day - to track one's progress -
```

Can you learn anything from daily tracking data other than when you did smth?

## Time eaters

-   where do you lose the most time?
-   

## Sleeping patterns

```{r}
df_activity_hours <- data_long %>%
   mutate(date = as.Date(datetime)) %>%
   group_by(date, activity) %>%
   summarise(hours = n() * (20/60)) %>%
   ungroup() 

# Filter the data for "Sleep" activity and remove the first and last rows
df_filtered_sleep <- df_activity_hours %>%
  filter(activity %in% "Sleep") %>%
  arrange(date) %>%
  slice(2:(n()-1))

# Create a new data frame that includes both the current and next row's data
df_segments <- df_filtered_sleep %>%
  mutate(date_next = lead(date),
         hours_next = lead(hours)) %>%
  filter(!is.na(date_next))

# Plot using geom_segment
ggplot() +
  geom_segment(data = df_segments, aes(x = date, y = hours, xend = date_next, yend = hours_next, 
                                       color = ifelse(hours < 7, "Less than 7", "More than 7")), size = 0.75) +
  scale_color_manual(values = c("Less than 7" = "red", "More than 7" = "blue")) +
  labs(title = "Sleeping patterns",
       x = "",
       y = "Hours") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_blank())
```

```{r}
library(tsibble) #allowing temporal structures in tidy data frames
library(tseries) #time series analysis
library(feasts) #tools for analyzing TS
library(fable) #models for TS
library(tibble) #creates tidy data frames






# Forecast

# sources: 
# https://rpubs.com/ReynaldiGev15/timeseries-forecasting
# https://bookdown.org/singh_pratap_tejendra/intro_time_series_r/
# https://otexts.com/fpp3/
```

```{r}
#Visualize the distribution of sleep hours by day of the week from data_long
data_long %>% 
  group_by(activity, day_of_week) %>%
  filter(activity %in% c("Sleep")) %>% 
  summarise(hours = n() * (20/60),
            mean = mean(hours)/52) %>%
  mutate(weekend = ifelse(day_of_week %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
         day_of_week = factor(day_of_week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  ungroup() %>% 
  ggplot(aes(x = day_of_week, y = mean, fill = weekend)) +
  geom_bar(stat = "identity") +
  geom_point(size = 7) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))
```



## Eating patterns

```{r}
data_long %>% 
  group_by(activity, day_of_week) %>%
  filter(activity %in% c("Food")) %>% 
  summarise(hours = n() * (20/60),
            mean = mean(hours)/52) %>%
  mutate(weekend = ifelse(day_of_week %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
         day_of_week = factor(day_of_week, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>% 
  ungroup() %>% 
  ggplot(aes(x = day_of_week, y = mean, fill = weekend)) +
  geom_bar(stat = "identity") +
  geom_point(size = 7) +
  labs(x = "",
       y = "Hours") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))
```

```{r}
data_long %>% 
  filter(activity %in% "Food") %>% 
  ggplot(aes(y=day_of_week, x=as.POSIXct(time_column, format="%H:%M"))) +  # Convert time_column to POSIXct
  geom_bin2d(position = "identity", bins = c(72, 7)) +
  scale_fill_gradient(low = "yellow", high = "orange") +
  scale_x_time(labels = scales::time_format("%H:%M"), 
               expand = c(0, 0),
               breaks = seq(from = as.POSIXct("00:00", format="%H:%M"), 
                            to = as.POSIXct("23:40", format="%H:%M"), 
                            by = "20 min")) +  # Add scale_y_time here
  coord_flip() +
  labs(y = "", x = "") +
  theme(axis.text.y = element_text(size = 4.5),
        axis.text.x = element_text(size = 10),
        axis.title = element_blank(),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

```{r}
df_filtered_food <- df_activity_hours %>% 
  filter(activity %in% "Food") %>%
  arrange(date) %>%
  slice(2:(n()-1)) %>% 
  mutate(rolling_mean = zoo::rollmean(hours, k = 7, fill = NA, align = "right"))

ggplot(df_filtered_food, aes(x = date)) +
  geom_line(aes(y = hours, color = "Food"), size = 0.5) +  # Original data
  geom_line(aes(y = rolling_mean, color = "Rolling Mean"), size = 1) +  # Rolling mean
  scale_x_date(limits = c(min(df_filtered_food$date), max(df_filtered_food$date)), 
                   expand = c(0, 0),
                   date_breaks = "1 month",  # Display every month on the x-axis
                   date_labels = "%B") +  # Format the date labels as "Month Year"
  labs(title = "How much time have been spent eating each day in 2020?",
       x = NULL,
       y = "",
       color = "Activity") +  # Labels
  theme(legend.position = "none",
        plot.title = element_text(family = "IBM Plex Mono", hjust = 0.5, size = 18),
        axis.text = element_text(family = "IBM Plex Mono"),
        axis.title = element_text(family = "IBM Plex Mono"),
        axis.text.x = element_text(size = 7)) +
  scale_color_manual(values = c("Food" = "gray", "Rolling Mean" = "#FFD700")) # Custom colors
```


```{r}
ggplot(data_long, aes(x = as.POSIXct(datetime), y = sleep)) +
  geom_line() + 
  facet_wrap(~month, nrow = 12)
```


## Behavioral patterns

-   you can't always lie "you can't lie consistently"


::: column-page-inset
```{r}
circle_plot <-
   data_long |>
   filter(activity == "Food") |>
   count(day, hour) |>
   mutate(
     hour_label = case_when(
       hour == 12 ~ "12pm",
       hour == 0 ~ "12am",
       hour > 12 ~ paste0(hour - 12, "pm"),
       hour < 12 ~ paste0(hour, "am")
     ),
     pct = n / sum(n)
     )
 
 
tooltip_hour <- function(day, hour_label, ...) {
   this_hour_count <-
     circle_plot |>
     filter(hour_label == !!hour_label)
 
   this_hour_total <- sum(this_hour_count$n)
   this_hour_pct <- scales::percent(this_hour_total / sum(circle_plot$n), 0.1)
   this_hour_total <- trimws(format(this_hour_total, big.mark = ","))
 
   this_hour_days <-
     this_hour_count |>
     mutate(
       across(pct, scales::percent_format(0.1)),
       across(n, format, big.mark = ","),
       across(n, trimws),
       text = glue("{day}: {pct} ({n})"),
       text = if_else(day == !!day, glue("<strong>{text}</strong>"), text)
     ) |>
     glue_data("<li>{text}</li>") |>
     glue_collapse()
 
   glue::glue(
     "<p><strong>{hour_label}</strong><br><small>{this_hour_pct} of total ({this_hour_total})</small></p>",
     "<ul>{this_hour_days}</ul>"
   )
}
 
circle_plot$tooltip <- pmap_chr(circle_plot, tooltip_hour)
 
 
plot_time_of_day <-
   ggplot(circle_plot) +
   aes(y = n, fill = day, x = hour, data_id = hour, tooltip = tooltip) +
   geom_area(
     data = function(d) {
       # Shade from midnight-6am and 6pm-midnight, kinda like geom_step_area()
       max_count <- max(d$n)
       tibble(
         day = sort(rep(unique(d$day), 6)),
         # hour = rep(c(0, 6, 6.01, 18, 18.01, 24), 7),
         hour = rep(c(6, 12, 12.01, 6, 6.01, 12), 7),
         n = rep(c(max_count, max_count, 0, 0, max_count, max_count), 7),
       tooltip = ""
       )
     },
     fill = "#aaaaaa30",
   ) +
   ggiraph::geom_col_interactive(show.legend = FALSE, width = 1) +
   facet_wrap(vars(day), nrow = 2) +
   coord_polar(start = pi) +
   scale_x_continuous(
   breaks = seq(0, 23, 3),
     minor_breaks = 0:23,
     labels = c("12am", paste0(seq(3, 9, 3), "am"), "12pm", paste0(seq(3, 9, 3), "pm")),
     limits = c(0, 24),
     expand = expansion()
   ) +
   scale_y_continuous(expand = expansion(), breaks = seq(0, 100, 25)) +
   scale_fill_discrete() +
   labs(
     title = "Did I really waste more time in the morning??",
     x = NULL,
     y = NULL
   ) +
 #  theme_classic() +
   theme(
     axis.text.y = element_blank(),
     axis.text.x = element_text(size = 10),
     panel.grid.major.y = element_blank()
)
 
ggiraph::girafe(
   ggobj = plot_time_of_day,
   width_svg = 12,
   height_svg = 8,
   options = list(
     ggiraph::opts_hover_inv("filter: saturate(30%) brightness(125%)"),
     ggiraph::opts_hover(css = "opacity:1"),
     ggiraph::opts_tooltip(
       placement = "container",
       css = "width: 12rem; font-family: var(--font-monospace, 'IBM Plex Mono');",
       # These don't matter, position is set by CSS rules below
       offx = 830,
       offy = 400,
       use_cursor_pos = FALSE
     )
   ),
   desc = knitr::opts_current$get("fig.alt")
)

#Code taken and adapted from Garrick Aden's blog post [Visualizing mytime](https://www.garrickadenbuie.com/blog/tweet-archive-in-r/).
```
:::


Why do I waste time?

(what comes before wasting time)

## Parting thoughts

Tracking one's daily life has many merits:

-   allows one to reflex in the moment if one has been doing what one wanted;

-   gives you visual confirmation of how you spend your time;

-   plan activities that can be achieved with some certainty when the $n$ hours has been spend on that activity

but it doesn't provide answer as to whether you are living your best life.

It can provide hints like having a routine, sleeping 7-8 hours a day, eating 3 meals a day, spending time communicating, socializing. But it can never tell you if that specific moment is significant or not.

On its own a single cell in the middle of your day tells nothing, but if you look at the pattern or lack thereof, you can see which areas of your life you might want to look into.


[Part 5](https://vyudenko.blog/posts/time-tracking-5)